Exploring the role of electric field in the catalase active site

### 2018-03-17 bb494/manuscript.tex
add efield figures
new stuff

An issue with mutagenesis studies like Chelikani \textit{et. al.} is that catalase's active site is very sensitive to changes. Mutagenesis of many residues results in large changes in enzyme kinetics \cite{substrateflow}. Here we conduct a less invasive experiment to probe the role of electric field in the catalase active site. We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (Figure \ref{fig:hypothesis}.e). This change is important because it a.) is tiny, adding only 30 $\AA^3$ of steric bulk \cite{3ntsize}, and b.) incorporates a hydroxy group of pKa 7.0 into the catalase active site \cite{3ntsize}. Catalase activity is pH-invariant in the range of 5 to 9 \cite{phdependence,kcatkm}; thus, by incorporating a group which gains charge at pH 7, we can precisely study the effect of charge in the catalase active site on enzyme kinetics without worrying about the undesired effects of mutagenesis.\\

\begin{figure}
  \begin{subfigure}{0.49\textwidth}
    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
    \begin{minipage}{.9\textwidth}
      %% \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
      \includegraphics[width=\linewidth]{figures/channel-visualization.png}
    \end{minipage}
  \end{subfigure}
  \begin{subfigure}{0.5\textwidth}
    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
    \begin{minipage}{.9\textwidth}
      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
      \includegraphics[width=0.5\linewidth]{figures/active-site-visualization.png}%
    \end{minipage}
  \end{subfigure}
  \caption{Depiction of catalase active pocket. \textbf{Left:} wildtype pocket with proposed ligand-orienting E-field \cite{electricpotential}. The field should orient H2O2 for optimal reaction with the heme. \textbf{Right:} modified HPII-3NY with a deprotonated 3-nitrotyrosine residue at site 206, with hypothetical augmented E-field. The disrupted E-field should result in un- or mis-aligned substrate.}
  \label{fig:hypothesis}
\end{figure}

@@

>>>>

An issue with mutagenesis studies like Chelikani \textit{et. al.} is that catalase's active site is very sensitive to changes. Replacing residues in the active site can dramatically impact enzyme efficiency \cite{substrateflow}. A better technique would purturb the electric field without full amino acid mutagenesis. Here we propose a new method for studying the catalase active site: genetic code expansion. We demonstrate how noncanonical amino acids can be incorporated into the catalase active site, as described in \textit{Hamill et. al.}, and how these modifications allow a less drastic change in active site environment. We use genetic code expansion \cite{hammill} to incorporate two functional groups onto F206 in the catalase active site, turning it into 3-nitrotyrosine (Figure \ref{fig:hypothesis}.e). This change is important because it a.) is tiny, adding only 30 $\AA^3$ of steric bulk \cite{3ntsize}, and b.) incorporates a hydroxy group of pKa 7.0 into the catalase active site \cite{3ntsize}. Catalase activity is pH-invariant in the range of 5 to 9 \cite{phdependence,kcatkm}; thus, by incorporating a group which gains charge at pH 7, we can precisely study the effect of charge in the catalase active site on enzyme kinetics without worrying about the undesired effects of mutagenesis.\\

\begin{figure}
  \begin{subfigure}{0.4\textwidth}
    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
    \begin{minipage}{.9\textwidth}
      %% \includegraphics[width=0.5\linewidth]{figures/cartoon-visualization.png}%
      \includegraphics[width=\linewidth]{figures/channel-visualization.png}
    \end{minipage}
  \end{subfigure}%
  \begin{subfigure}{0.6\textwidth}
    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
    \begin{minipage}{.9\textwidth}
      \includegraphics[width=0.4\linewidth]{figures/efield-wt.pdf}%
      \hspace{1cm}
      \includegraphics[width=0.4\linewidth]{figures/efield-3ny.pdf}%
    \end{minipage}
  \end{subfigure}
  \vspace{3cm}
  \begin{subfigure}{0.5\textwidth}
    \begin{minipage}{0.1\textwidth}\caption{}\end{minipage}%
    \begin{minipage}{.9\textwidth}
      \includegraphics[width=\linewidth]{figures/3ny.png}%
      \end{minipage}
  \end{subfigure}
  \caption{Depiction of catalase active pocket. \textbf{Left:} wildtype pocket with proposed ligand-orienting E-field \cite{electricpotential}. The field should orient H2O2 for optimal reaction with the heme. \textbf{Right:} modified HPII-3NY with a deprotonated 3-nitrotyrosine residue at site 206, with hypothetical augmented E-field. The disrupted E-field should result in un- or mis-aligned substrate.}
  \label{fig:hypothesis}
\end{figure}

@@

ncAA substitution is a powerful technique which allows easy testing of hypotheses in regions which are very sensitive to structural change. This technique could be used to test -other-catalase-hypotheses-.\\

Perhaps this technique could be used to study e-fields generally.\\

We propose that the following residues be incorporated into the channel at sites - - -. Coupled with APBS simulations, these results could be a powerful test of the effects of catalase e-field.\\
